Multi-level generalizations of collective spin dynamics and squeezing

### 2019-10-10 collective_sun/sun_notes.tex
simplified work in drive operator product appendix
removing item from to-do list
fixed error in drive operator product appendix
minor simplifications and additions to drive operator product section

\begin{itemize}
\item Products and commutators of drive operators
\item Visualization with generalized Bloch spheres
  \begin{itemize}
  \item Single-particle eigenstates of drive operators
  \item Single-particle dynamics induced by drive operators
  \end{itemize}
\item Classifying collective states (e.g.~eigenstates of interactions,
  spin-wave-like states)
\item Implications of SOC
  \begin{itemize}
  \item Effective collective dynamics for $\abs{\phi}\ll1$
  \item Deriving \eqref{eq:A_L_phi_small}
  \item Special angles $\phi$
  \end{itemize}
\item Generalizations of squeezing beyond SU(2)?
\item Consider different interactions
  \begin{itemize}
  \item Include electronic states
  \item Super-exchange regime
  \item Non-uniform (``disordered'') interactions
  \end{itemize}
\item Consider different drives
\end{itemize}

@@

The drive operators are defined by
\begin{align}
  D_{L,0} \equiv T_{L,0},
  &&
  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
  &&
  \sigma_M \equiv \sign\p{M}^{3/2},
\end{align}
and can alternately be expressed as
\begin{align}
  D_{LM} = \f{\tau_M}{\sqrt{2}} \p{T_{LM} + \epsilon_M T_{L,-M}},
\end{align}
where
\begin{align}
  \tau_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0, \\
    \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
  &&
  \epsilon_M \equiv \p{-1}^M \sign\p{M},
\end{align}
which also implies that
\begin{align}
  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
  &&
  \tilde\tau_M \equiv
  \begin{cases}
    e^{-i\pi/4} & M = 0 \\
    \sign\p{M}^{M-1/2} & M \ne 0
  \end{cases}.
\end{align}
In order to evaluate the product of two drive operators, we can thus
expand them in terms of transition operators, evaluate the
corresponding products of transition operators (see Appendix
\ref{sec:trans_prod}), and finally expand the result in terms of drive
operators.  The general product of two drive operators takes the form
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
  \sum_L \p{f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
    + f_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
\end{align}
with operators
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1+m_2}}
  \sp{\tilde\tau_{m_1+m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}},
  \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1-m_2}}
  \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}.
\end{align}
The general product of drive operators simplifies somewhat in several
special cases.  If one of $m_1,m_2$ is zero, for example, then
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  \stackrel{0\in\set{m_1,m_2}}{=}
  \sum_L f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2} \times
  \begin{cases}
    D_{L,m_1+m_2} & L + \ell_1 + \ell_2 ~\t{even} \\
    i D_{L,-\p{m_1+m_2}} & \t{otherwise}
  \end{cases}.
\end{align}
If $\abs{m_1}=\abs{m_2}$, then
\begin{multline}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  \stackrel{\abs{m_1}=\abs{m_2}}{=}
  \sqrt{s_m} \p{-1}^m \sum_L f_{\ell_1 m;\ell_2,-m}^{L,0} D_{L,0} \times
  \begin{cases}
    1 & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \\
    0 & \t{otherwise}
  \end{cases} \\
  + \f{\sqrt{\sign\p{m}}}{\sqrt{s_m}} \f1{\sqrt{2}}
  \sum_L f_{\ell_1 m;\ell_2 m}^{L,2m} \times
  \begin{cases}
    D_{L,2m} & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \sign\p{m} \\
    i D_{L,-2m} & \t{otherwise}
  \end{cases},
\end{multline}
where $m \equiv m_1$ and $s_m\equiv+1$ ($-1$) if $m_1=m_2$
($m_1\ne m_2$).  Finally, if $\abs{m_1}\ne\abs{m_2}$ with
$m_1,m_2\ne0$, then
\begin{multline}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  \stackrel{\substack{\abs{m_1}\ne\abs{m_2}\\m_1,m_2\ne0}}{=}
  \f1{\sqrt{2}} \sum_{s\in\set{\pm1}}\f1{\sqrt{s}}
  \tau_{m_1} \tau_{s m_2} \tau_{m_1+s m_2}
  \sign\p{m_1+sm_2} \\
  \times \sum_L f_{\ell_1 m_1;\ell_2 sm_2}^{L,m_1+sm_2} \times
  \begin{cases}
    D_{L,m_1+sm_2} & \p{-1}^L = \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)} \\
    i D_{L,-\p{m_1+sm_2}} & \t{otherwise}
  \end{cases},
\end{multline}
where
\begin{align}
  \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)}
  \equiv s \p{-1}^{\ell_1+\ell_2}
  \sign\p{m_1} \sign\p{sm_2} \sign\p{m_1+sm_2}.
\end{align}

>>>>

\begin{itemize}
\item Visualization with generalized Bloch spheres
  \begin{itemize}
  \item Single-particle eigenstates of drive operators
  \item Single-particle dynamics induced by drive operators
  \end{itemize}
\item Classifying collective states (e.g.~eigenstates of interactions,
  spin-wave-like states)
\item Implications of SOC
  \begin{itemize}
  \item Effective collective dynamics for $\abs{\phi}\ll1$
  \item Deriving \eqref{eq:A_L_phi_small}
  \item Special angles $\phi$
  \end{itemize}
\item Generalizations of squeezing beyond SU(2)?
\item Consider different interactions
  \begin{itemize}
  \item Include electronic states
  \item Super-exchange regime
  \item Non-uniform (``disordered'') interactions
  \end{itemize}
\item Consider different drives
\end{itemize}

@@

The drive operators are defined by
\begin{align}
  D_{L,0} \equiv T_{L,0},
  &&
  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
  &&
  \sigma_M \equiv \sign\p{M}^{3/2},
\end{align}
and can alternately be expressed as
\begin{align}
  D_{LM} = \f{\tau_M}{\sqrt{2}} \p{T_{LM} + \epsilon_M T_{L,-M}},
\end{align}
where
\begin{align}
  \tau_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0, \\
    \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
  &&
  \epsilon_M \equiv \p{-1}^M \sign\p{M},
\end{align}
which also implies that
\begin{align}
  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
  &&
  \tilde\tau_M \equiv
  \begin{cases}
    \sqrt{2}/\p{1+i} & M = 0 \\
    \sign\p{M}^{M-1/2} & M \ne 0
  \end{cases}.
\end{align}
In order to evaluate the product of two drive operators, we can thus
expand them in terms of transition operators, evaluate the
corresponding products of transition operators (see Appendix
\ref{sec:trans_prod}), and finally expand the result in terms of drive
operators.  The general product of two drive operators takes the form
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
  \sum_L \p{f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
    + f_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
\end{align}
with operators
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\equiv \sum_{s\in\set{\pm1}}
  \f12 \sp{\tilde\tau_{m_1+m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}}
  \sqrt{s} D_{L,s\p{m_1+m_2}},
  \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\equiv \sum_{s\in\set{\pm1}}
  \f12 \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}
  \sqrt{s} D_{L,s\p{m_1-m_2}}.
\end{align}
If $m_1\pm m_2=0$, these operators simplify to
\begin{align}
  E_{\ell_1 m;\ell_2,-m}^{(L,+)}
  &= \f1{\sqrt{2}} \sp{1 - \p{-1}^{L+\ell_1+\ell_2}
    \p{1-\delta_{m,0}}} D_{L,0}, \\
  E_{\ell_1 m;\ell_2 m}^{(L,-)}
  &= \f{\epsilon_m}{\sqrt{2}}
  \sp{1 + \p{-1}^{L+\ell_1+\ell_2}} D_{L,0},
\end{align}
and are otherwise
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\stackrel{m_1+m_2\ne0}{=}
  \tau_{m_1+m_2}^* \sum_{s\in\set{\pm1}}
  \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \epsilon_{m_1+m_2}}
  \sqrt{s} D_{L,s\p{m_1+m_2}}, \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\stackrel{m_1-m_2\ne0}{=}
  \tau_{m_1-m_2}^* \sum_{s\in\set{\pm1}}
  \f12 \sp{\epsilon_{m_2} + s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_1-m_2}}
  \sqrt{s} D_{L,s\p{m_1-m_2}}.
\end{align}
If $\p{m_1,m_2}=\p{0,m}$, then we can use the fact that
$i\tau_m\tau_{-m}^*=\epsilon_m$ for $m\ne0$ to simplify
\begin{align}
  D_{\ell_1,0} D_{\ell_2 m}
  = \sum_L f_{\ell_1,0,\ell_2,m}^{Lm}
  \sum_{s\in\set{\pm1}} \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}}
  \sqrt{s} D_{L,sm}.
\end{align}
